*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-16 Code/Manuscript_O2.Rmd
Added results and discussion

# PURPARFits <- readRDS(file.path(DataIn, "PICO_Merged_GrowthFitsPURPAR.Rds"))

@@

## *Prochlorococcus* evolutionary and functional diversity

@@

## *Prochlorococcus* responses to changing niches

@@

We then analyzed growth rates in terms of cumulative diel Photosynthetically Usable Radiation (PUR). Photosynthetic organisms rely on absorbing light energy within the PAR range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. PUR represents the fraction of PAR that is absorbed by the pigments of photosynthetic organisms [@morelAvailableUsableStored1978]. PUR, therefore, takes into account the pigment composition of photosynthetic organisms and the specific spectral wavebands these pigments absorb.  *Prochlorococcus marinus* light-harvesting complexes, [@chisholmProchlorococcusMarinusNov1992;
@goerickePigmentsProchlorococcusMarinus1992], show an absorption maxima
of 442 nm for divinyl chlorophyll a and 478 nm for divinyl chlorophyll b
[@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to
efficiently harvestblue light in the 400 nm
to 500 nm range [@morelAvailableUsableStored1978] prevailing in deep ocean
habitats, where only blue spectral wavelengths prevail [@holtropVibrationalModesWater2021].  In *Prochlorococcus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993]. Given the different spectral light regimes typical of the niches of different ecotypes,  expressing growth rates in terms of cumulative diel absorbed PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel absorbed PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth.

@@

## Photosystem II maintenance as a limitation on *Prochlorococcus* growth

@@

Three xenic cultures of *P. marinus* were obtained from Bigelow Labs, NCMA
Maine, US (REFERENCE). MED4 (CCMP1986) is from High Light-adapted (HLI) clade; SS120
(CCMP1375) is from Low Light-adapted (LLIII) clade; and MIT9313 (CCMP2773)
is from Low Light-adapted (LLIV) clade. Cultures were
maintained in incubators set to 22°C with a light/dark cycle of 12 h, with illumination from XXXXX. The PAR level for maintenance cultures reflected  Photosynthetically Active Radiation (PAR) in the source
niche of the ecotype; MED4, of 160 µmol photons m^-2^ s^-1^; SS120 and
MIT9313 at 30 µmol photons m^-2^ s^-1^. To maintain active growth all strains were
transferred weekly with XX to YY dilution with Pro99 media
[@mooreCulturingMarineCyanobacterium2007] prepared with autoclaved artificial
seawater (National
Center for Marine Algae and Microbiota (NCMA) protocol by combining salt
solution I and salt solution II using the enriched artificial seawater
(ESAW) recipe.  REPLACe with CITATION OF NCMA protocol)XXXX

Controlled growth experiments were performed using PSI Multicultivators (Figure \@ref(fig:LabeledMC); MCMIX-OD orMC1000-OD, Brno, Czech Republic ).  Each multicultivator individually controls 8 tubes at a common temperature of 22°C.  Each tube containing 70 mL of Pro99 media was inoculated with 10 mL of growing maintenance culture. In a factorial matrix design, each tube was then subject to an individual combination of sinusoidal photoperiod (4, 8, 12, 16  h);  reaching a peak PAR (30, 90, 180 µmol photons m^-2^ s^-1^), with defined spectral bandwidth (AA, BB, CC, DD nm).  [O~2~] levels (2.5 µM, 25 µM, 250 µM) were imposed by bubbling tubes with varying ratios of air and
Nitrogen (N~2~), with consistent 0.05% of Carbon Dioxide (CO~2~)
gas, delivered through a 0.2 μm sterile microfilter via a G400 gas mixing system XXXXGIVE SOURCEXXXX. [O~2~] in situ was verified using oxygen optodes (PyroScience,
Germany) inserted into tubes for
real-time measurements, with a temperature probe in
the aquarium of the bioreactor to correct [O~2~] measures for temperature
fluctuations. In addition, the Pyroscience software corrected [O~2~] based on the
salinity of the media (32 ppt). The flow rate of the gas mixture was controlled, but variations in bubbling speed, PAR and culture density affected the [O~2~] achieved in each tube. A low [O~2~] of 0.5 µM - 5 µM (reported as 2.5 µM hereafter),  was achieved by sparging with a gas mixture containing 99.95% N~2~ and 0.05%
CO~2~. An intermediate [O~2~] of 10 - 25 µM (reported hereafter as 25 µM) was achieved by sparging with a gas mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. A high O~2~ of 200 µM - 280 µM (reported hereafter as 250 µM) was achieved by sparging with lab air (78% N~2~, 21% O~2~, 1% Ar and 0.05% CO~2~).

@@

In situ XXXXItalicsXXX measurements of Optical Density (OD) 680 nm, a proxy for cell suspension density, cell size dependent scatter and cell chlorophyll content;  and OD 720 nm, a proxy for cell suspension density and cell size dependent scatter, were recorded every 5 minutes over least 8 to 14 days,  depending on the
duration of the lag phase, if any. All data from the Multicultivator were saved
as comma separated values file  (CITATION of FILES LOCATION).

PAR of 180, 90 or 30 µmol photons m^-2^ s^-1^, and spectral wavebands (white LED
full spectrum, 660 nm, 530 nm, and 450 nm) were chosen to approximate light
levels and spectral colors spanning the vertical ocean water column, from
near-surface to the lower euphotic zone depths. Photoperiods were chosen to approximate diel cycles characteristic of current and hypothetical future
niches of *P. marinus*; 16 h represents
temperate (45°N) summer at the ocean surface; 12 h for equatorial (0°N)
ocean surface or temperate (45°N) spring and fall ocean surface or
temperate (45°N) summer at deeper ocean depths; 8 h for temperate (45°N)
winter at the surface or at temperate (45°N) spring and fall at depth
and equatorial (0°N) deep ocean depths; and 4 h for temperate (45°N)
winter or deep ocean depths during temperate (45°N) spring and fall.

@@

### Estimation of Photosynthetically Usable Radiation

XXXXTEXT NEEDEDXXXX
```{r PurParPlot, fig.cap = PurParPlot_cap , fig.height = 20, fig.width = 10, echo = FALSE}

@@

Furthermore,
Evidence of proteins
derived from HL ecotypes of *P. marinus* inOMZ at depths up to 200
meters,  with O~2~ of 15 µM. The extent to which [O~2~] defines
the niches occupied by different *P. marinus* ecotypes, as compared to
potentially covarying environmental variables like photoperiod, light
spectrum, and light level, is poorly described.

@@

@@

@@

@@

@@

@@

@@

>>>>

 PURPARFits <- readRDS(file.path(DataIn, "PICO_Merged_GrowthFitsPURPAR.Rds"))

@@

## *Prochlorococcus* evolutionary and functional diversity {.unnumbered}

@@

## *Prochlorococcus* responses to changing niches {.unnumbered}

@@

We then analyzed growth rates in terms of cumulative diel Photosynthetically Usable Radiation (PUR). Photosynthetic organisms rely on absorbing light energy within the PAR range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. PUR represents the fraction of PAR that is absorbed by the pigments of photosynthetic organisms [@morelAvailableUsableStored1978]. PUR, therefore, takes into account the pigment composition of photosynthetic organisms and the specific spectral wavebands these pigments absorb.  *Prochlorococcus marinus* light-harvesting complexes, [@chisholmProchlorococcusMarinusNov1992;
@goerickePigmentsProchlorococcusMarinus1992], show an absorption maxima
of 442 nm for divinyl chlorophyll a and 478 nm for divinyl chlorophyll b
[@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to
efficiently harvest blue light in the 400 nm
to 500 nm range [@morelAvailableUsableStored1978] prevailing in deep ocean
habitats, where only blue spectral wavelengths prevail [@holtropVibrationalModesWater2021].  In *Prochlorococcus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993]. Given the different spectral light regimes typical of the niches of different ecotypes,  expressing growth rates in terms of cumulative diel absorbed PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel absorbed PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth.

@@

## Photosystem II maintenance as a limitation on *Prochlorococcus* growth {.unnumbered}

@@

Three xenic cultures of *P. marinus* were obtained from Bigelow Labs, NCMA
Maine, US (REFERENCE). MED4 (CCMP1986) is from High Light-adapted (HLI) clade; SS120
(CCMP1375) is from Low Light-adapted (LLIII) clade; and MIT9313 (CCMP2773)
is from Low Light-adapted (LLIV) clade. Cultures were
maintained in incubators set to 22°C with a light/dark cycle of 12 h, with illumination from XXXXX. The PAR level for maintenance cultures reflected  Photosynthetically Active Radiation (PAR) in the source
niche of the ecotype; MED4, of 160 µmol photons m^-2^ s^-1^; SS120 and
MIT9313 at 30 µmol photons m^-2^ s^-1^. To maintain active growth all strains were
transferred weekly with 1 to 5 dilution with Pro99 media
[@mooreCulturingMarineCyanobacterium2007] prepared with autoclaved artificial
seawater (National
Center for Marine Algae and Microbiota (NCMA) protocol by combining salt
solution I and salt solution II using the enriched artificial seawater
(ESAW) recipe.  REPLACe with CITATION OF NCMA protocol)XXXX

Controlled growth experiments were performed using PSI Multicultivators (Figure \@ref(fig:LabeledMC); MCMIX-OD or MC1000-OD, Brno, Czech Republic ).  Each multicultivator individually controls 8 tubes at a common temperature of 22°C.  Each tube containing 70 mL of Pro99 media was inoculated with 10 mL of growing maintenance culture. In a factorial matrix design, each tube was then subject to an individual combination of sinusoidal photoperiod (4, 8, 12, 16  h);  reaching a peak PAR (30, 90, 180 µmol photons m^-2^ s^-1^), with defined spectral bandwidth (AA, BB, CC, DD nm).  [O~2~] levels (2.5 µM, 25 µM, 250 µM) were imposed by bubbling tubes with varying ratios of air and
Nitrogen (N~2~), with consistent 0.05% of Carbon Dioxide (CO~2~)
gas, delivered through a 0.2 μm sterile microfilter via a G400 gas mixing system XXXXGIVE SOURCEXXXX. [O~2~] in situ was verified using oxygen optodes (PyroScience,
Germany) inserted into tubes for
real-time measurements, with a temperature probe in
the aquarium of the bioreactor to correct [O~2~] measures for temperature
fluctuations. In addition, the Pyroscience software corrected [O~2~] based on the
salinity of the media (32 ppt). The flow rate of the gas mixture was controlled, but variations in bubbling speed, PAR and culture density affected the [O~2~] achieved in each tube. A low [O~2~] of 0.5 µM - 5 µM (reported as 2.5 µM hereafter),  was achieved by sparging with a gas mixture containing 99.95% N~2~ and 0.05%
CO~2~. An intermediate [O~2~] of 10 - 25 µM (reported hereafter as 25 µM) was achieved by sparging with a gas mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. A high O~2~ of 200 µM - 280 µM (reported hereafter as 250 µM) was achieved by sparging with lab air (78% N~2~, 21% O~2~, 1% Ar and 0.05% CO~2~).

@@

*In situ* measurements of Optical Density (OD) 680 nm, a proxy for cell suspension density, cell size dependent scatter and cell chlorophyll content;  and OD 720 nm, a proxy for cell suspension density and cell size dependent scatter, were recorded every 5 minutes over least 8 to 14 days,  depending on the
duration of the lag phase, if any. All data from the Multicultivator were saved
as comma separated values file  (CITATION of FILES LOCATION).

PAR of 180, 90 or 30 µmol photons m^-2^ s^-1^, and spectral wavebands (white LED
full spectrum, 660 nm, and 450 nm) were chosen to approximate light
levels and spectral colors spanning the vertical ocean water column, from
near-surface to the lower euphotic zone depths. Photoperiods were chosen to approximate diel cycles characteristic of current and hypothetical future
niches of *P. marinus*; 16 h represents
temperate (45°N) summer at the ocean surface; 12 h for equatorial (0°N)
ocean surface or temperate (45°N) spring and fall ocean surface or
temperate (45°N) summer at deeper ocean depths; 8 h for temperate (45°N)
winter at the surface or at temperate (45°N) spring and fall at depth
and equatorial (0°N) deep ocean depths; and 4 h for temperate (45°N)
winter or deep ocean depths during temperate (45°N) spring and fall.

@@

### Estimation of Photosynthetically Usable Radiation {.unnumbered}

To calculate the Photosynthetic Usable Radiation (PUR), a proxy of photons absorbed, for each ecotype
of *P. marinus*, the imposed Photosynthetic Active
Radiation (PAR) was first determined. An emission profile of each colored LED
light of the MCMIX-OD Multicultivator and the white LED light of the MC1000-OD Multicultivator was obtained using a Jaz spectrometer (Ocean Optics, Inc.,Dunedin FL, USA) equipped with a fiber
optic cable, HH2 FiberOpticJmp (Part number A901073, Malvern Panalytical
Ltd, Malvern, UK) and scanned to measure the emission wavelengths from 200 nm to 1000 nm. Following the determination of PAR, an *in-vivo* whole cell absorbance spectrum for each *P. marinus* was obtained using the Olis 14 UV/VIS Clarity
Spectrophotometer (Olis Inc, Bogart, Georgia) by scanning across the visible spectral range of λ = 350 nm to 750 nm at 1 nm intervals. A whole cell absorbance spectrum was measured for each strain under each imposed spectral waveband. The 20 cm path length of the internally reflective cavity of the Olis spectrophotometer was corrected to a 1 cm path length using the Javorfi correction method [@javorfiQuantitativeSpectrophotometryUsing2006] on a PRO 99 media subtracted whole cell absorbance spectra of each strain.

A weighted PUR spectrum was determined by applying an integrated weighted function to the blank subtracted, whole cell absorbance spectrum for each *P. marinus* ecotype and the emission spectrum of the imposed LED growth light [@morelAvailableUsableStored1978]. The whole cell absorbance spectrum was normalized to the absorbance maximum of divinylchlorophyll a (Chl a~2~), determined dynamically between 400 nm and 460 nm, and each LED spectrum was normalized to the emission maximum.

Equation \@ref(eq:WeightedPUR) was used to determine the weighted PUR spectrum P(λ) according to [@morelAvailableUsableStored1978],

\begin{equation}
  P(λ) = A(λ) × E(λ) 
  (\#eq:WeightedPUR)
\end{equation}

where P(λ) is the weighted PUR absorbance spectrum over 400 nm to 700 nm spectral range, A(λ) is the Chl a~2~ normalized whole cell absorbance spectrum over 400 nm to 700 nm and E(λ) is the imposed growth light emission spectrum over 400 nm to 700 nm.

Actinic absorbed PUR levels (µmol photons m^-2^ s^-1^) were calculated from actinic PAR (µmol photons m^-2^ s^-1^) levels using the equation \@ref(eq:PUR) from [@morelAvailableUsableStored1978],

\begin{equation}
  PUR = \frac{\int_{400}^{700} P(λ)}{\int_{400}^{700} E(λ)} × PAR
  (\#eq:PUR)
\end{equation}

where P(λ) is the weighted PUR absorbance spectrum from equation \@ref(eq:WeightedPUR), E(λ) is the imposed growth light emission spectrum from equation \@ref(eq:WeightedPUR) and PAR is the imposed light level (µmol photons m^-2^ s^-1^).  Figure (\@ref(fig:PurParPlot)) shows the calculated absorbed PUR (µmol photons m^-2^ d^-1^) versus imposed actinic PAR (µmol photons m^-2^ s^-1^) for each strain and each spectral waveband (nm).

The applied photoperiod was administered using a sinusoidal function to simulate light levels equivalent to sun rise and sunset. The area under a sine curve is equivalent to a triangle, therefore the equation to determine the cumulative diel absorbed PUR (µmol photons m^-2^ d^-1^) is one half of the base (photoperiod) multiplied by the height (PUR) equation \@ref(eq:SumPURDay).

\begin{equation}
  Cumulative diel absorbed PUR = \frac{PUR × 3600 × Photoperiod}{2}
  (\#eq:SumPURDay)
\end{equation}

where PUR is the actinic absorbed light calculated from equation \@ref(eq:PUR) (µmol photons m^-2^ s^-1^), 3600 is the time conversion from seconds to hour and photoperiod is the imposed photoperiod (h).

```{r PurParPlot, fig.cap = PurParPlot_cap , fig.height = 20, fig.width = 10, echo = FALSE}

@@

Furthermore,
Evidence of proteins
derived from HL ecotypes of *P. marinus* in OMZ at depths up to 200
meters,  with O~2~ of 15 µM. The extent to which [O~2~] defines
the niches occupied by different *P. marinus* ecotypes, as compared to
potentially covarying environmental variables like photoperiod, light
spectrum, and light level, is poorly described.

@@

### MED4 PAR and growth trends {.unnumbered}

#### High oxygen environment {.unnumbered}

The growth rate of MED4, from the high light (HLI) clade of *P.
marinus*, increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), when grown under 250 µM O~2~. No growth was
observed under any imposed conditions under a 4 h photoperiod. MED4 was
able to grow under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ but only when the blue light photoperiod was longer than 4 h; the
white LED light photoperiod was longer than 8 h; or when the red light
photoperiod was longer than 16 h. Growth rate increased
with increasing photoperiods for all spectral wavebands when exposed to
PAR of 90 and 180 µmol photons m^-2^ s^-1^. MED4 grew fastest for each
spectral waveband under 16 h photoperiods. The maximum growth rate
(µ~max~) was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 180 uE blue
light and 16 h photoperiod.

#### Intermediate oxygen environment {.unnumbered}

Similar to the growth trends of the 250 µM O~2~ experiments, MED4
maintained at 25 µM O~2~, showed the fastest growth rate when the
photoperiod was 16 h for each spectral waveband, across PAR levels
(Figure \@ref(fig:MED4PAR)). The 25 µM O~2~ growth experiments showed
generally similar growth rate as the 250 µM O~2~ growth experiments,
with the interesting exception that growth rates increased by 40% under
25 µM O~2~ at PAR of 30 µmol photons m^-2^ s^-1^ and 16 h photoperiod,
and even showed growth under 12 h red light photoperiod, in contrast to
the 250 µM O~2~ growth experiments. White LED light treatments grew
slower, if at all, under 25 µM O~2~, only growing under higher PAR of 90
and 180 µmol photons m^-2^ s^-1^ and longer photoperiods of 12 and 16 h.
As under 250 µM O~2~, the fastest growth rate observed for each spectral
waveband at each imposed PAR were greatest under the 16 h photoperiod.
The µ~max~ was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "25", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^  achieved under 180 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod. The 4 h photoperiod
experiments under white LED light were not performed because no growth
was achieved when grown under an 8 h photoperiod.

#### Low oxygen environment {.unnumbered}

MED4 did not grow under any tested conditions sparged to the lowest O~2~
of 2.5 µM (Figure \@ref(fig:MED4PAR)). 2.5 µM O~2~ growth experiments
were not conducted for 4 and 16 h photoperiods, as no growth occurred
when MED4 was exposed to 8 and 12 h photoperiods.

@@

### SS120 PAR and growth trends {.unnumbered}

#### High oxygen environment {.unnumbered}

The growth rate of SS120, from the low light (LLIII) clade of *P.
marinus*, increased with lower imposed PAR and longer photoperiods, when
grown under 250 µM O~2~ (Figure \@ref(fig:SS120PAR)). SS120 was able to
grow under the lowest PAR treatment of 30 µmol photons m^-2^ s^-1^ but
only when the photoperiod was longer than 4 h under blue, white, and red
light. SS120 required longer than 8 h photoperiod of imposed green
light. No growth was observed under any blue light photoperiods when
exposed to PAR of 90 µmol photons m^-2^ s^-1^. Growth rate increased
with increasing photoperiods for white, red and green light when exposed
to PAR of 90 µmol photons m^-2^ s^-1^; however, no growth was observed
when the red light photoperiod was 16 h. Experiments conducted under PAR
of 180 µmol photons m^-2^ s^-1^ showed no growth under imposed blue and
red lights. Growth rate decreased with longer photoperiods and showed
growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol
photons m^-2^ s^-1^ and white LED light. The growth rate generally
increased with longer photoperiods under green light treatments, however
no growth was observed under the 8 h photoperiod. The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^  achieved under 90 µmol photons
m^-2^ s^-1^ white LED light and 16 h photoperiod.

#### Intermediate oxygen environment {.unnumbered}

Similar to the growth trends of the 250 µM O~2~ and PAR of 30 µmol
photons m^-2^ s^-1^ experiments, SS120 maintained at 25 µM O~2~, no
growth was observed under a 4 h photoperiod for imposed red spectral
waveband, however under blue light, SS120 was able to grow (Figure
\@ref(fig:SS120PAR)). In contrast to the growth trends of the 250 µM
O~2~ and PAR of 90 µmol photons m^-2^ s^-1^ experiments, SS120 grew
under 4 and 8 h blue light and 16 h red light photoperiods. The growth
rate decreased with increasing photoperiod under PAR of 90 µmol photons
m^-2^ s^-1^ white LED light treatments. Blue light treatments under PAR
of 180 µmol photons m^-2^ s^-1^ showed growth only under an 8 h
photoperiod. This growth rate decreased by 53% compared to the 8 h blue
light photoperiod under PAR of 90 µmol photons m^-2^ s^-1^. The µ~max~
was
`r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ (Table. \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons
m^-2^ s^-1^ blue light and 8 h photoperiod.

The 25 µM O~2~, less than 16 h photoperiod and 180 µmol photons m^-2^ s^-1^ experiments under white LED
light were not performed due to time constraints.

#### Low oxygen environment {.unnumbered}

SS120 only grew
(`r PURPARFits %>% filter(Strain == "SS120", WL == "660", O2_uM == "2.5", Photoperiod == 12) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^) under PAR of 30 µmol photons m^-2^ s^-1^ and 12 h red light
photoperiod when sparged to the lowest O~2~ of 2.5 µM (Figure
\@ref(fig:SS120PAR)). SS120 did not grow under any other tested
conditions. 2.5 µM O~2~ growth experiments were not conducted for 4 and
16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^, as no
growth occurred when SS120 was exposed to 8 and 12 h photoperiods. Red
light 16 h photoperiod experiments were not performed due to time
constraints.

@@

\newpage
### MIT9313 PAR results {.unnumbered}

#### High oxygen environment {.unnumbered}

The growth rate of MIT9313, from the low light (LLIV) clade of *P.
marinus*, increased with low to moderate imposed PAR and longer
photoperiods, when grown under 250 µM O~2~ (Figure
\@ref(fig:MIT9313PAR)). Growth rate increased with increasing
photoperiods under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ for blue, white LED and red light. The green light treatment
showed growth only when the photoperiod was longer than 12 h and
decreased by \~ 50% under a 16 h photoperiod. Under PAR of 90 µmol
photons m^-2^ s^-1^, growth rate increased with longer photoperiods for
white LED, red, and green light treatments. Growth rates decreased from
4 to 8 h blue light photoperiod and exhibited no growth under
photoperiods of 12 and 16 h. Blue light did not induce growth at 180
µmol photons m^-2^ s^-1^ under any photoperiod; white LED and red light
treatments only exhibited growth under the 8 h photoperiod and growth
rates increased with longer photoperiods under green light. The µ~max~
was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 30 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod.

#### Intermediate oxygen environment {.unnumbered}

Similar to the growth trends of the 250 µM O~2~ and PAR of 30 µmol
photons m^-2^ s^-1^ experiments, MIT9313 maintained at 25 µM O~2~,
growth rate increased with increasing photoperiods for all spectral
wavebands tested (Figure \@ref(fig:MIT9313PAR)). The 25 µM O~2~ growth
experiments showed generally similar growth rate as the 250 µM O~2~ and
PAR of 30 µmol photons m^-2^ s^-1^, with the interesting exception that
the growth rate increased by 50% under 25 µM O~2~ and 16 h blue light
photoperiod. In contrast to the blue light growth responses under 250 µM
O~2~ and PAR of 90 µmol photons m^-2^ s^-1^, growth rate increased with
longer photoperiods and even grew under 12 and 16 h photoperiods. In
fact, the fastest growth rate of all the O~2~ experiments was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "25", Photoperiod == 16, Par_ue == 90) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 16 h blue light
photoperiod and PAR of 90 µmol photons m^-2^ s^-1^ under 25 µM O~2~.
Growth rate increased by 50% under the 4 h blue light photoperiod; by
400% under the 8 h blue light photoperiod; by 40% under the 12 h red
light photoperiod and by 80% under the 16 h white LED photoperiod,
compared to the 250 µM O~2~ and PAR of 90 µmol photons m^-2^ s^-1^
growth experiments. In contrast to the 250 µM O~2~ growth experiments,
MIT9313 was able to grow when exposed to PAR of 180 µmol photons m^-2^
s^-1^ and blue light under all photoperiods except 16 h; additionally,
white LED and red light treatments induced growth across all tested
photoperiods under 25 µM O~2~.

The 25 µM O~2~, 4 h photoperiod experiments under white LED light and
all green light were not performed due to time constraints.

#### Low oxygen environment {.unnumbered}

The growth experiments under 2.5 µM O~2~ for MIT9313 showed growth
responses for all spectral wavebands except white LED light, albeit
slower than under the 25 µM O~2~ experiments (Figure
\@ref(fig:MIT9313PAR)). MIT9313 was able to grow under PAR of 30 µmol
photons m^-2^ s^-1^ but only under an 8 h blue and red light
photoperiod. Growth rate increased with longer photoperiods under blue
light treatments and PAR of 90 µmol photons m^-2^ s^-1^ however, growth
was inhibited under a 16 h blue light photoperiod and only grew under an
8 h photoperiod under red and green light treatments. The growth rate
for MIT9313 under PAR of 180 µmol photons m^-2^ s^-1^ and blue light
treatment decreased with increasing photoperiods with growth inhibition
under a 16 h photoperiod. The red light treatments under PAR of 180 µmol
photons m^-2^ s^-1^ showed similar growth rate for 8 and 12 h
photoperiods and no growth were observed for green light treatments. The
µ~max~ was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "2.5", Photoperiod == 4, Par_ue == 180) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 4 h blue light
photoperiod and PAR of 180 µmol photons m^-2^ s^-1^.

The 2.5 µM O~2~ white LED treatments under 4, 8 and 16 h photoperiods
and red light under 4 and 16 h photoperiods were not performed due to
time constraints.

@@

#### Generalized Additive Model (GAM) of MED4 growth {.unnumbered}

XXX Fig letters do not match as now we have red GAMs. Below is results for Blue GAMsXXXX

Due to the complex interactions of light level and photoperiod, a GAM
was employed (Figure \@ref(fig:MED4GAM)) to model the growth trends
of MED4 in response to changing blue and red light PAR and photoperiod, across
the 3 different imposed oxygen concentrations. Under 250 µM O~2~, Figure
\@ref(fig:MED4GAM)A showed that the fastest growth rates were
achieved above 90 µmol photons m^-2^ s^-1^ and photoperiods greater than
12 h, indicated by the contour line labeled 0.62 d^-1^ (representing the
90^th^ percentile of achieved growth rate). Figure
\@ref(fig:MED4GAM)A also illustrated that MED4 exploited all light
levels as long as the photoperiods were longer than 4 h, albeit at
slower growth rate at lower PAR and shorter photoperiods. Under 25 µM
O~2~, MED4 grew, but required longer than 4 h photoperiod (Figure
\@ref(fig:MED4GAM)B). Interestingly, the GAM illustrated no
difference in growth rate when MED4 was grown under 25 µM or 250 µM O~2~
with photoperiods greater than 12 h. Figure \@ref(fig:MED4GAM)C
showed that MED4 did not grow under the tested conditions at 2.5 µM
O~2~.

@@

#### Generalized Additive Model (GAM) of SS120 growth {.unnumbered}

XXX Fig letters do not match as now we have red GAMs. Below is results for Blue GAMsXXXX

The GAM model in Figure \@ref(fig:SS120GAM) displays the growth
trends of SS120 in response to changing blue light PAR and photoperiod,
across the 3 different imposed oxygen concentrations. Under 250 µM O~2~,
Figure \@ref(fig:SS120GAM)A illustrated that the fastest growth
rates were achieved below PAR of 50 µmol photons m^-2^ s^-1^ and
photoperiods between 8 and 12 h, indicated by the contour line labeled
0.29 d^-1^ (representing the 90^th^ percentile of achieved growth rate).
Figure \@ref(fig:SS120GAM)A also showed that SS120 does not grow at
PAR above 90 µmol photons m^-2^ s^-1^ and the growth rate are slower as
photoperiods go above 12 or below 8 h. Under 25 µM O~2~, SS120 grew
under a photoperiod of 8 h and exploited all light levels and grew under
a 12 h photoperiod but at PAR of 50 µmol photons m^-2^ s^-1^. The
fastest growth rate achieved was at a higher PAR of \~100 µmol photons
m^-2^ s^-1^, indicated by the contour line labeled 0.5 d^-1^
(representing the 90^th^ percentile of achieved growth rate), out pacing
the 90^th^ percentile fastest growth rates under 250 µM O~2~. Figure
\@ref(fig:SS120GAM)C showed that SS120 did not grow under the tested
conditions at 2.5 µM O~2~.

@@

#### Generalized Additive Model (GAM) of MIT9313 growth {.unnumbered}

XXX Fig letters do not match as now we have red GAMs. Below is results for Blue GAMsXXXX

The GAM model in Figure \@ref(fig:MIT9313GAM) displays the growth
trends of MIT9313 in response to changing blue light PAR and
photoperiod, across the 3 different imposed oxygen concentrations. Under
250 µM O~2~, Figure \@ref(fig:MIT9313GAM)A illustrated that the
fastest growth rates were achieved between PAR of 30 µmol photons m^-2^
s^-1^ and 50 µmol photons m^-2^ s^-1^ and photoperiods longer than 8 h,
indicated by the contour line labeled 0.51 d^-1^ (representing the
90^th^ percentile of achieved growth rates). Figure
\@ref(fig:MIT9313GAM)A also illustrated that growth rate increases
with longer photoperiods, as long as the PAR levels are below 50 µmol
photons m^-2^ s^-1^ and that MIT9313 does not grow when exposed to PAR
levels above 90 µmol photons m^-2^ s^-1^. In contrast to the 250 µM O~2~
growth experiment, MIT9313 was able to exploit all actinic PAR levels
(Figure \@ref(fig:MIT9313GAM)B). Interestingly, the growth rate was
similar across all PAR levels between 4 and 12 h photoperiods, indicated
by the contour line labeled 0.29 d^-1^ (representing the 10^th^
percentile of achieved growth rate) and 0.44 d^-1^ (representing the
50^th^ percentile of achieved growth rate) (Figure
\@ref(fig:MIT9313GAM)B). However, for photoperiods exceeding 12 h,
the growth rate showed a decline at higher PAR levels but showed the
overall fastest growth rate achieved was at PAR levels less than 100
µmol photons m^-2^ s^-1^, indicated by the contour line labeled 0.69
d^-1^ (representing the 90^th^ percentile of achieved growth rate)
(Figure \@ref(fig:MIT9313GAM)B). Figure \@ref(fig:MIT9313GAM)C
showed that MIT9313 is capable of growing under 2.5 µM O~2~ as long as
the photoperiod is between 4 and 8 h and between 50 to 150 µmol photons
m^-2^ s^-1^ PAR levels (Figure \@ref(fig:MIT9313GAM)C).

@@

## Discussion PAR and growth trends {.unnumbered}

The purpose of this study was to determine the interactive effects of
photoperiod, PAR level, spectral bandwidth and dissolved [O~2~] regimes
on the growth of *P. marinus*, under simulated present day and
hypothetical future ecological niches.

### Photoperiod {.unnumbered}

The representative of HLI clade, *P. marinus* MED4, was completely
excluded from growth under the 4 h photoperiod, and then showed
increasing growth rate with increasing photoperiod. Thus, MED4 can
potentially exploit longer summer photoperiods in temperate zones, but
will remain excluded from growth in winter through its requirement for
more than 4 h of light per day.

The LLIII clade, *P. marinus* SS120, grew under photoperiods longer than
4 h and showed increasing growth rate with increasing photoperiods.
Therefore, SS120 can potentially exploit the longer summer photoperiods
in temperate zones but cannot sustain growth in winter through its
requirement for more than 4 h of light per day.

The LLIV clade, *P. marinus* MIT9313, exploited all tested photoperiods,
and showed increasing growth rate with increasing photoperiods. Hence,
MIT9313 can potentially inhabit temperate zones with shorter winter and
longer summer photoperiods, and exploit habitats at depth subject to
photoperiod attenuation.

### PAR Levels {.unnumbered}
